---
title: "p8105_hw5_rx2236"
author: "Ruiqi Xue"
date: "2023-11-12"
output: github_document
---


```{r}
library(tibble)
library(tidyverse)
library(purrr)
library(broom)
library(ggplot2)
set.seed(1)
```

### Problem 2

Create a dataframe and read in the data iteratively.
```{r}
p2_df = 
  tibble(
    file_name = list.files("data/")
    ) |>
  mutate(data = map(str_c("data/", file_name), read_csv)) |>
  unnest()
  
```

Tidy the dataframe.
```{r}
p2_df = p2_df |>
  mutate(file_name = str_replace(file_name, ".csv", ""),
         arm = str_sub(file_name, 1, 3),
         id = str_sub(file_name, 5, 7)) |>
  pivot_longer(week_1:week_8,
               names_to = "week",
               values_to = "observation",
               names_prefix = "week_") |>
  mutate(week = as.numeric(week),
         arm = case_match(
           arm,
           "con" ~ "control",
           "exp" ~ "experimental"
         )) |>
  select(arm, id, week, observation)

p2_df
  
```

Make a spaghetti plot showing observations on each subject over time.

```{r}
p2_df |>
  group_by(arm, id) |>
  ggplot(aes(x = week, y = observation, color = id)) +
  geom_line(alpha = 0.8) +
  facet_grid(~arm) +
  geom_smooth(aes(group = 1), se = FALSE, color = "black") +
  labs(x = 'Week', y = 'Observation', title = 'Plot of Observations on Subjects in Each Arm Over Time')
```

Compare the two arms, we see that values in the control arm fluctuate placidly over time, whereas values in the experimental arm are higher than the ones in the control arm in general, and have an increasing tendency over time.






